\textbf{Proof Reconstruction in Classical Propositional Logic}

### 2017-04-27 slides/slides.tex
[ slides ] minor changes.

\setmathfont[ExternalLocation=fonts/
  , BoldFont=xits-mathbold.otf
  ]{xits-math.otf}

@@

\usepackage[autostyle]{csquotes}
\usepackage[
    backend=biber
  , style=authoryear-icomp
  , sortlocale=en_US
  , natbib=true
  , url=false
  , doi=true
  , eprint=false
]{biblatex}
\addbibresource{ref.bib}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}

@@

\begin{frame}[plain]
\titlepage
% \begin{tikzpicture}[overlay,remember picture, scale=0.5]
% \foreach \x in {0,4,..., \textwidth}
%   \foreach \y in {0,1,..., 20} {
%     \node[opacity=0.05](tptp) at (\x,\y){$Γ ⊢ ϕ$};
%   }
% \end{tikzpicture}
\end{frame}

% \begin{frame}
%   \frametitle{Outline}
%   \tableofcontents
% \end{frame}

@@

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{It is available at \url{http://www.gilith.com/software/metis}}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{It is available at \url{http://github.com/jonaprieto/athena}}
  };
}
\only<6->{
\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{It is available at \url{http://github.com/agda/agda}}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<8->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

\only<2->{ \draw[->, thick] (problem) to (tptp)};
\only<3->{ \draw[->, thick] (tptp) to (metis)};
\only<4->{ \draw[->, thick] (metis) to (tstp)};
\only<5->{ \draw[->, thick] (tstp) to (athena)};
\only<6->{ \draw[->, thick] (athena) to (agdafile)};
\only<7->{ \draw[->, thick] (agdafile) to (agda)};
\only<8->{ \draw[->, thick] (agda) to (verified)};
\only<8->{ \draw[->, thick, gray] (agda) to (failure)};
\end{tikzpicture}
\vfill
\end{frame}

@@

\begin{lrbox}{\agdapragma}
\begin{lstlisting}
$ cat Or.agda
module Or where

data _or_ (A B : Set) : Set where
  inj1 : A -> A or B
  inj2 : B -> A or B

postulate
  A B    : Set
  or-comm : A or B -> B or A
{-# ATP prove or-comm #-}
\end{lstlisting}
\end{lrbox}

\begin{tikzpicture}[scale=2]
\only<1->{ \node(agda-pragmas) at (0,0){\includegraphics{figures/agda-pragmas}}};
\only<1>{
  \node[fill=aliceblue, rectangle](atp-pragma) at (3,0){
  \scalebox{0.7}{\usebox\agdapragma}
  }};

\only<2->{ \node(eagda) at (2,0){
  \includegraphics{figures/eagda}
  \footnote{Development version of Agda in order to handle a new built-in ATP-pragma.
    \url{https://github.com/asr/eagda}}
  }};

\only<3->{ \node(apia) at (4,0){
  \includegraphics{figures/apia}
  \footnote{Haskell program for proving first-order theorems written in Agda using ATPs.
    \url{https://github.com/asr/apia}}}
  };

\only<3->{ \node(tptp) at (2,-2){\includegraphics{figures/tptp}}};

\only<3->{ \node(atp) at (4,-2){\includegraphics{figures/atp}}};

\only<2->{\draw[->,ultra thick] (agda-pragmas) to (eagda)};

\only<3->{\draw[->,ultra thick] (eagda) to (apia)};
\only<3->{\draw[->, very thick, dashed, gray] (apia) to (tptp)};
\only<3->{\draw[->, very thick, dashed, gray] (tptp) to (atp)};
\only<3->{\draw[->, very thick, dashed, gray] (atp) to (apia)};

\end{tikzpicture}

\end{frame}

@@

  \begin{itemize}
  \item Is a language\footnote{Is available at
        \url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}}
        to encode problems
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF %THF, TFF,
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture, among others
      \item[\texttt{formula}] version in TPTTP format
    \end{itemize}
  \end{itemize}

@@

\begin{frame}[fragile, label=tptp-examples]{TPTP Examples}{From Prop-Pack repository
  \footnote{Is available at \url{http://github.com/jonaprieto/pro-pack}}}

@@

\begin{frame}[fragile, label=metis-rules]{Inference Rules of Metis}
\hyperlink{tstp-example}{TSTP} derivations exhibit these inferences:
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|cl| }
\hline
\textbf{Rule} & \textbf{Task} \\
\hline
\hyperlink{atp-canonicalize}{\texttt{canonicalize}} & transforms formulas to CNF, DNF or NNF\\
\hyperlink{atp-clausify}{\texttt{clausify}} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & extracts a formula from a conjunction\\
\hyperlink{atp-negate}{\texttt{negate}} & applies negation to the formula\\
\hyperlink{atp-resolve}{\texttt{resolve}} & applies theorems of resolution\\
\hyperlink{atp-simplify}{\texttt{simplify}} & applies over a list of formula to simplify them\\
\hyperlink{atp-strip}{\texttt{strip}} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=atp-canonicalize]{Canonicalize}{Inference Rule}
\begin{block}{canonicalize}

\end{block}
\end{frame}

\begin{frame}[fragile, label=atp-clausify]{Clausify}{Inference Rule}
\begin{block}{clausify}

\end{block}
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{Conjunct}{Inference Rule}
\begin{block}{conjunct}
\begin{lstlisting}
$ cat ~/agda-metis/src/ATP/Metis/Rules/Conjunct.agda
...
conjunct : Prop -> Prop -> Prop
conjunct (φ ∧ ψ) ω with ⌊ eq φ ω ⌋ | ⌊ eq ψ ω ⌋
... | true  | _     = φ
... | false | true  = ψ
... | false | false = conjunct φ ω
conjunct φ ω        = φ

atp-conjunct
  : ∀ {Γ} {φ}
  -> (ω : Prop)
  -> Γ ⊢ φ
  -> Γ ⊢ conjunct φ ω
\end{lstlisting}
\end{block}

\end{frame}

\begin{frame}[fragile, label=atp-negate]{Negate}{Inference Rule}
\begin{block}{negate}

\end{block}
\end{frame}

\begin{frame}[fragile, label=atp-resolve]{Resolve}{Inference Rule}
\begin{block}{resolve}

\end{block}
\end{frame}

\begin{frame}[fragile, label=atp-simplify]{Simplify}{Inference Rule}
\begin{block}{simplify}

\end{block}
\end{frame}

\begin{frame}[fragile, label=atp-strip]{Strip}{Inference Rule}
\begin{block}{strip}
\end{block}
\end{frame}

@@

>>>>

\setmathfont[ExternalLocation=fonts/
  , BoldFont=xits-mathbold.otf
  ]{xits-math.otf}
\newfontfamily\mathfont{fonts/xits-math.otf}

@@

\usepackage[autostyle]{csquotes}
\usepackage[
    backend=biber
  , style=authoryear-icomp
  , sortlocale=en_US
  , natbib=true
  , url=false
  , doi=true
  , eprint=false
]{biblatex}
\addbibresource{ref.bib}
\renewcommand*{\nameyeardelim}{\addcomma\addspace}
\usepackage{silence}
\WarningFilter{biblatex}{Patching footnotes failed}
\WarningFilter{hyperref}{Token not allowed in a PDF string}

@@

\begin{frame}[plain]
\titlepage
  \begin{tikzpicture}[overlay, remember picture] (eafit) at (current page.center)
    \node[xshift=0cm,yshift=0cm]
      {\includegraphics[width=0.25\textwidth]{figures/eafit}};
    \node[xshift=3cm,yshift=-2cm] (chalmers) at (current page.center)
      {\includegraphics[width=0.25\textwidth]{figures/chalmers}};
  \end{tikzpicture}
\end{frame}

@@

\only<2->{\node[right = 0.8cm of problem, hyperlink node=tptp-syntax](tptp){
\includegraphics[scale=0.7]{figures/tptp}
};
}
\only<3->{
\node[right= 0.8cm of tptp, hyperlink node=metis] (metis)
  {\includegraphics[scale=0.7]{figures/metis}\,
  \footnote{It is available at \url{http://www.gilith.com/software/metis}}
  };
}
\only<4->{
\node[right= 0.8cm of metis, hyperlink node=tstp-syntax] (tstp)
  {\includegraphics[scale=0.7]{figures/tstp}};
}
\only<5->{
\node[below= 1.5cm of tstp, hyperlink node=athena] (athena)
  {\includegraphics[scale=0.7]{figures/athena}
  \footnote{It is available at \url{http://github.com/jonaprieto/athena}}
  };
}
\only<6->{
\node[left = 0.5cm of athena, hyperlink node=agda-file] (agdafile)
  {\includegraphics[scale=0.7]{figures/agdafile}
  };
}
\only<7->{
\node[left = 0.5cm of agdafile, hyperlink node=agda] (agda)
  {\includegraphics[scale=0.7]{figures/agda}
  \footnote{It is available at \url{http://github.com/agda/agda}}
  };
}
\only<8->{
\node[below = 0.18cm of problem, hyperlink node=verified-example] (verified)
  {\includegraphics[scale=0.7]{figures/verified}};
}
\only<9->{
\node[below = 0.18cm of verified, opacity=0.5, hyperlink node=failure-example] (failure)
  {\includegraphics[scale=0.7]{figures/failure}};
}

\only<2->{ \draw[->, thick] (problem) to (tptp)};
\only<3->{ \draw[->, thick] (tptp) to (metis)};
\only<4->{ \draw[->, thick] (metis) to (tstp)};
\only<5->{ \draw[->, thick] (tstp) to (athena)};
\only<6->{ \draw[->, thick] (athena) to (agdafile)};
\only<7->{ \draw[->, thick] (agdafile) to (agda)};
\only<8->{ \draw[->, thick] (agda) to (verified)};
\only<9->{ \draw[->, thick, gray] (agda) to (failure)};
\end{tikzpicture}
\vfill
\end{frame}

@@

@@

  \begin{itemize}
  \item Is a language\footnote{Is available at
        \url{http://www.cs.miami.edu/~tptp/TPTP/SyntaxBNF.html}}
        to encode problems \citep{Sut09}
  \item Is the input of the ATPs
  \item Annotated formulas with the form
   \begin{center}
\begin{tptp}
language(name, role, formula).
\end{tptp}
    \end{center}
    \begin{itemize}
      \item[\texttt{language}] FOF or CNF %THF, TFF,
      \item[\texttt{name}] to identify the formula within the problem
      \item[\texttt{role}] axiom, definition, hypothesis, conjecture, among others
      \item[\texttt{formula}] version in TPTTP format
    \end{itemize}
  \end{itemize}

@@

\begin{frame}[fragile, label=tptp-examples]{TPTP Examples}{From Prop-Pack repository
  \footnote{Is available at \url{http://github.com/jonaprieto/prop-pack}}}

@@

\begin{frame}[fragile, label=metis-rules]{Inference Rules of Metis}
\hyperlink{tstp-example}{TSTP} derivations exhibit these inferences:
\vfill
\begin{table}[!ht]
\begin{center}
\scalebox{0.8}{
{\renewcommand{\arraystretch}{1.3}%
\begin{tabular}{|cl| }
\hline
\textbf{Rule} & \textbf{Task} \\
\hline
\texttt{canonicalize} & transforms formulas to CNF, DNF or NNF\\
\texttt{clausify} & performs clausification\\
\hyperlink{atp-conjunct}{\texttt{conjunct}} & extracts a formula from a conjunction\\
\texttt{negate} & applies negation to the formula\\
\texttt{resolve} & applies theorems of resolution\\
\texttt{simplify} & applies over a list of formula to simplify them\\
\texttt{strip} & splits a formula into subgoals\\[1ex]
\hline
\end{tabular}}
}
\end{center}
\label{tab:label}
\end{table}
\vfill
\gobackreconstruction
\end{frame}

\begin{frame}[fragile, label=atp-conjunct]{Conjunct}{Inference Rule}
\begin{block}{conjunct}
\begin{agda}
conjunct : Prop → Prop → Prop
conjunct (φ ∧ ψ) ω with ⌊ eq φ ω ⌋ | ⌊ eq ψ ω ⌋
... | true  | _     = φ
... | false | true  = ψ
... | false | false = conjunct φ ω
conjunct φ ω        = φ

atp-conjunct
  : ∀ {Γ} {φ}
  → (ω : Prop)
  → Γ ⊢ φ
  → Γ ⊢ conjunct φ ω
\end{agda}
\end{block}
\end{frame}

@@

\begin{lrbox}{\agdapragma}
\begin{agda}
module Or where

data _OR_ (A B : Set) : Set where
  inj1 : A → A OR B
  inj2 : B → A OR B

postulate
  A B    : Set
  OR-comm : A OR B → B OR A
{-# ATP prove OR-comm #-}
\end{agda}
\end{lrbox}

\begin{tikzpicture}
\only<1->{ \node(agdafile) at (0,0){
  \includegraphics[scale=0.8]{figures/agda-pragmas}}};
\only<1>{
  \node[fill=aliceblue, rectangle, right = 2.5cm of agdafile] (atp-pragmas){
  \scalebox{0.7}{\usebox\agdapragma}
  }
  };

\only<2->{
  \node[right=1cm of agdafile] (eagda){
  \includegraphics[scale=0.8]{figures/eagda}
  \footnote{Development version of Agda in order to handle a new built-in ATP-pragma.
    \url{https://github.com/asr/eagda}}
  }
};

\only<2->{
\node[right=1cm of eagda] (agdai){
  \includegraphics[scale=0.8]{figures/agdai}
  }
};

\only<3->{
  \node[below=0.4cm of agdai] (apia){
  \includegraphics[scale=0.8]{figures/apia}
  \footnote{Haskell program for proving first-order theorems written in Agda using ATPs.
    \url{https://github.com/asr/apia}}}
  };

\only<4->{
  \node[left= 1cm of apia] (tptp)
    {\includegraphics[scale=0.8]{figures/tptp}}};

\only<5->{ \node[below = 0.8cm of apia] (atp)
  {\includegraphics[scale=0.8]{figures/atp}}};

\only<7->{ \node[below = 2.4 of agdafile, rectangle, text=white,fill=black, align=left, font=\tiny] (proved)
{
\texttt{\$ agda Or.agda}\\
\texttt{\$ apia Or.agda --atp=online-metis}\\
\texttt{Metis---2.3 proved the conjecture}
  }
};

\only<2->{\draw[->, ultra thick] (agdafile) to (eagda)};
\only<2->{\draw[->, ultra thick] (eagda) to (agdai)};
\only<3->{\draw[->, ultra thick] (agdai) to (apia)};
\only<4->{\draw[->, very thick, dashed, gray] (apia) to (tptp)};
\only<5->{\draw[->, very thick, dashed, gray] (tptp) to (atp)};
\only<6->{\draw[->, very thick, dashed, gray] (atp) to (apia)};

\end{tikzpicture}
\end{frame}
